---
title: "Projet Multi-omics Single Cell"
author: "Sreevatsa and Akshaya"
date: "2024-10-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Multi-Omics Single Analysis

## Quality Control

Load Library:

```{r loading}
# Setting up environment ===================================================

# Clean environment
rm(list=ls())
rm(list = ls(all.names = TRUE)) # will clear all objects including hidden objects
gc() # free up memory and report the memory usage
options(max.print = .Machine$integer.max, scipen = 999, stringsAsFactors = F, dplyr.summarise.inform = F) # avoid truncated output in R console and scientific notation


library(Seurat)
library(ggplot2)
library(dplyr)
library(Matrix)


path <-"C:/Users/sreev/Desktop/Cours/ING5 Verona/PROJET SingleCell R/TAF/NSCLC_donor_7_count_sample_feature_bc_matrix/sample_feature_bc_matrix"
sparse_matrix <- Seurat::Read10X(data.dir = path)

```

```{r separate}

seu <- Seurat::CreateSeuratObject(counts = sparse_matrix$`Gene Expression`,
                                  project = "NSCLC",
                                  min.cells = 3,
                                  min.features = 100)
```

```{r , echo=FALSE}
Seurat::FeatureScatter(seu, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") + geom_smooth(method = "lm")
Seurat::VlnPlot(seu, features = c("nCount_RNA",
                                  "nFeature_RNA")) 
```

```{r, echo=FALSE}

# Calculer les quartiles, min, max, VAMA et VAMI pour chaque groupe
stats <- VlnPlot(seu, features = c("nFeature_RNA"), pt.size = 0)$data %>%
  group_by(ident) %>%
  summarise(Min = min(nFeature_RNA),
            Q1 = quantile(nFeature_RNA, 0.25),
            Q2 = quantile(nFeature_RNA, 0.50),
            Q3 = quantile(nFeature_RNA, 0.75),
            Max = max(nFeature_RNA),
            VAMI = Q1 - 1.5 * (Q3 - Q1),  # Valeur adjacente minimale
            VAMA = Q3 + 1.5 * (Q3 - Q1))  # Valeur adjacente maximale

# Ajouter des lignes pour les quartiles, min, max, VAMI et VAMA sur le plot
VlnPlot(seu, features = c("nFeature_RNA"), pt.size = 0) + 
  geom_hline(data = stats, aes(yintercept = Min), linetype="dashed", color="purple") + 
  geom_hline(data = stats, aes(yintercept = Q1), linetype="dashed", color="blue") + 
  geom_hline(data = stats, aes(yintercept = Q2), linetype="dashed", color="green") + 
  geom_hline(data = stats, aes(yintercept = Q3), linetype="dashed", color="red") + 
  geom_hline(data = stats, aes(yintercept = Max), linetype="dashed", color="orange") + 
  geom_hline(data = stats, aes(yintercept = VAMI), linetype="dotted", color="darkblue") + 
  geom_hline(data = stats, aes(yintercept = VAMA), linetype="dotted", color="darkred") + 
  geom_text(data = stats, aes(y = Min, label = round(Min, 2)), vjust = -0.5, color="purple") +
  geom_text(data = stats, aes(y = Q1, label = round(Q1, 2)), vjust = -0.5, color="blue") +
  geom_text(data = stats, aes(y = Q2, label = round(Q2, 2)), vjust = -0.5, color="green") +
  geom_text(data = stats, aes(y = Q3, label = round(Q3, 2)), vjust = -0.5, color="red") +
  geom_text(data = stats, aes(y = Max, label = round(Max, 2)), vjust = -0.5, color="orange") +
  geom_text(data = stats, aes(y = VAMI, label = round(VAMI, 2)), vjust = -0.5, color="darkblue") +
  geom_text(data = stats, aes(y = VAMA, label = round(VAMA, 2)), vjust = -0.5, color="darkred")

```

```{r, echo=FALSE}
# mitochondrial genes
seu <- Seurat::PercentageFeatureSet(seu, 
                                    pattern = "^MT-", 
                                    col.name = "percent.mito")
# ribosomal genes
seu <- Seurat::PercentageFeatureSet(seu, 
                                    pattern = "^RP[SL]",
                                    col.name = "percent.ribo")
# hemoglobin genes (but not HBP)
seu <- Seurat::PercentageFeatureSet(seu,
                                    pattern = "^HB[^(P)]",
                                    col.name = "percent.globin")
Seurat::VlnPlot(seu, features = c("percent.mito",
                                  "percent.ribo",
                                  "percent.globin"))
```

```{r, echo=FALSE}

# Calculer les quartiles, min, max, VAMA et VAMI pour chaque groupe
stats <- VlnPlot(seu, features = c("percent.mito"), pt.size = 0)$data %>%
  group_by(ident) %>%
  summarise(Min = min(percent.mito),
            Q1 = quantile(percent.mito, 0.25),
            Q2 = quantile(percent.mito, 0.50),
            Q3 = quantile(percent.mito, 0.75),
            Max = max(percent.mito),
            VAMI = Q1 - 1.5 * (Q3 - Q1),  # Valeur adjacente minimale
            VAMA = Q3 + 1.5 * (Q3 - Q1))  # Valeur adjacente maximale

# Ajouter des lignes pour les quartiles, min, max, VAMI et VAMA sur le plot
VlnPlot(seu, features = c("percent.mito"), pt.size = 0) + 
  geom_hline(data = stats, aes(yintercept = Min), linetype="dashed", color="purple") + 
  geom_hline(data = stats, aes(yintercept = Q1), linetype="dashed", color="blue") + 
  geom_hline(data = stats, aes(yintercept = Q2), linetype="dashed", color="green") + 
  geom_hline(data = stats, aes(yintercept = Q3), linetype="dashed", color="red") + 
  geom_hline(data = stats, aes(yintercept = Max), linetype="dashed", color="orange") + 
  geom_hline(data = stats, aes(yintercept = VAMI), linetype="dotted", color="darkblue") + 
  geom_hline(data = stats, aes(yintercept = VAMA), linetype="dotted", color="darkred") + 
  geom_text(data = stats, aes(y = Min, label = round(Min, 2)), vjust = -0.5, color="purple") +
  geom_text(data = stats, aes(y = Q1, label = round(Q1, 2)), vjust = -0.5, color="blue") +
  geom_text(data = stats, aes(y = Q2, label = round(Q2, 2)), vjust = -0.5, color="green") +
  geom_text(data = stats, aes(y = Q3, label = round(Q3, 2)), vjust = -0.5, color="red") +
  geom_text(data = stats, aes(y = Max, label = round(Max, 2)), vjust = -0.5, color="orange") +
  geom_text(data = stats, aes(y = VAMI, label = round(VAMI, 2)), vjust = -0.5, color="darkblue") +
  geom_text(data = stats, aes(y = VAMA, label = round(VAMA, 2)), vjust = -0.5, color="darkred")

```

```{r, echo=FALSE}

# Calculer les quartiles, min, max, VAMA et VAMI pour chaque groupe
stats <- VlnPlot(seu, features = c("percent.ribo"), pt.size = 0)$data %>%
  group_by(ident) %>%
  summarise(Min = min(percent.ribo),
            Q1 = quantile(percent.ribo, 0.25),
            Q2 = quantile(percent.ribo, 0.50),
            Q3 = quantile(percent.ribo, 0.75),
            Max = max(percent.ribo),
            VAMI = Q1 - 1.5 * (Q3 - Q1),  # Valeur adjacente minimale
            VAMA = Q3 + 1.5 * (Q3 - Q1))  # Valeur adjacente maximale

# Ajouter des lignes pour les quartiles, min, max, VAMI et VAMA sur le plot
VlnPlot(seu, features = c("percent.ribo"), pt.size = 0) + 
  geom_hline(data = stats, aes(yintercept = Min), linetype="dashed", color="purple") + 
  geom_hline(data = stats, aes(yintercept = Q1), linetype="dashed", color="blue") + 
  geom_hline(data = stats, aes(yintercept = Q2), linetype="dashed", color="green") + 
  geom_hline(data = stats, aes(yintercept = Q3), linetype="dashed", color="red") + 
  geom_hline(data = stats, aes(yintercept = Max), linetype="dashed", color="orange") + 
  geom_hline(data = stats, aes(yintercept = VAMI), linetype="dotted", color="darkblue") + 
  geom_hline(data = stats, aes(yintercept = VAMA), linetype="dotted", color="darkred") + 
  geom_text(data = stats, aes(y = Min, label = round(Min, 2)), vjust = -0.5, color="purple") +
  geom_text(data = stats, aes(y = Q1, label = round(Q1, 2)), vjust = -0.5, color="blue") +
  geom_text(data = stats, aes(y = Q2, label = round(Q2, 2)), vjust = -0.5, color="green") +
  geom_text(data = stats, aes(y = Q3, label = round(Q3, 2)), vjust = -0.5, color="red") +
  geom_text(data = stats, aes(y = Max, label = round(Max, 2)), vjust = -0.5, color="orange") +
  geom_text(data = stats, aes(y = VAMI, label = round(VAMI, 2)), vjust = -0.5, color="darkblue") +
  geom_text(data = stats, aes(y = VAMA, label = round(VAMA, 2)), vjust = -0.5, color="darkred")

```

```{r clean seu, echo=FALSE}
seuClean <- subset(seu, subset = nFeature_RNA > 172 & nFeature_RNA < 5470 & percent.mito < 14 & percent.ribo > 5)

Seurat::VlnPlot(seuClean, features = c("nFeature_RNA","percent.mito", "percent.ribo"))
```

```{r, echo=FALSE}

most_expressed_boxplot <- function(object, ngenes = 20){
  
  # matrix of raw counts
  cts <- Seurat::GetAssayData(object, assay = "RNA", slot = "counts")
  
  # get percentage/cell
  cts <- t(cts)/colSums(cts)*100
  medians <- apply(cts, 2, median)
  
  # get top n genes
  most_expressed <- order(medians, decreasing = T)[ngenes:1]
  most_exp_matrix <- as.matrix((cts[,most_expressed]))
  
  # prepare for plotting
  most_exp_df <- stack(as.data.frame(most_exp_matrix))
  colnames(most_exp_df) <- c("perc_total", "gene")
  
  # boxplot with ggplot2
  boxplot <- ggplot(most_exp_df, aes(x=gene, y=perc_total)) +
    geom_boxplot() +
    coord_flip()
  return(boxplot)
}
most_expressed_boxplot(seu, 20)
most_expressed_boxplot(seuClean, 20)
```

## Normalization

```{r, echo=FALSE}
seuClean <- Seurat::NormalizeData(seuClean,
                     normalization.method = "LogNormalize",
                     scale.factor = 10000)
```

```{r, echo=FALSE}
seuClean <- Seurat::FindVariableFeatures(seuClean,
                            selection.method = "vst",
                            nfeatures = 2000) #CHANGE STV
```

```{r, echo=FALSE}
vf_plot <- Seurat::VariableFeaturePlot(seuClean)
top10 <- head(Seurat::VariableFeatures(seuClean), 10)
top10
Seurat::LabelPoints(plot = vf_plot,
            points = top10, repel = TRUE)
```

```{r, echo=FALSE}
seuClean <- Seurat::ScaleData(seuClean,
                 features = rownames(seuClean))
```

## Dimesionality Reduction

```{r, echo=FALSE}
seuClean <- Seurat::RunPCA(seuClean)
Seurat::DimPlot(seuClean, reduction = "pca") #  allows you to color the points in the plot according to a grouping variable (e.g. sample)
Seurat::FeaturePlot(seuClean, reduction = "pca", features = "nFeature_RNA") #allows you to color the points according to a continuous variable (e.g. gene expression)
```

```{r, echo=FALSE}
Seurat::FeaturePlot(seuClean, reduction = "pca", features = "MS4A1") #B cells markers
Seurat::FeaturePlot(seuClean, reduction = "pca", features = "COL3A1") #high variable markers
```

```{r, echo=FALSE}
Seurat::DimHeatmap(seuClean, dims = 1:12, cells = 500, balanced = TRUE)
Seurat::DimHeatmap(seuClean, dims = 13:20, cells = 500, balanced = TRUE)
```

```{r, echo=FALSE}
# Charger les bibliothèques nécessaires
library(Seurat)
library(ggplot2)

# Extraire la matrice de chargements pour chaque gène et chaque composant principal
pca_loadings <- seuClean[["pca"]]@feature.loadings

# Créer un graphique pour montrer l'impact des 10 gènes ayant les plus fortes contributions pour chaque composant
for (i in 1:20) {
  # Prendre les 10 valeurs les plus fortes en valeur absolue, qu'elles soient positives ou négatives
  top_genes_indices <- order(abs(pca_loadings[, i]), decreasing = TRUE)[1:10]
  top_genes_values <- pca_loadings[top_genes_indices, i]
  
  # Créer un barplot avec des couleurs en fonction des valeurs positives ou négatives
  barplot(top_genes_values,
          main = paste("Impact linéaire - PC", i),
          las = 2, col = ifelse(top_genes_values > 0, "blue", "red"),
          ylab = "Contribution",
          names.arg = rownames(pca_loadings)[top_genes_indices])
}


```

```{r, echo=FALSE}
for (i in 1:20) {
  print(VizDimLoadings(seuClean, dims = i, reduction = "pca"))
}

```

```{r, echo=FALSE}
 # vers PC_22 ça donne 2.5 de std

# Ajouter une ligne rouge à 2.5% d'écart type sur le plot
Seurat::ElbowPlot(seuClean, ndims = 40) + 
  geom_hline(yintercept = 5, color = "red", linetype = "dashed") +
  geom_text(aes(x = 13, y = 5, label = round(2.5, 2)), label="5% cut-in threshold", color = "red", vjust = -1, ) +
geom_hline(yintercept = 2.5, color = "green", linetype = "dashed") +
  geom_text(aes(x = 23, y = 3, label = round(2.5, 2)), label="2.5% cut-in threshold", color = "green", vjust = -1, )
```

```{r , echo=FALSE}
seuClean <- Seurat::RunUMAP(seuClean, dims = 1:6)
Seurat::DimPlot(seuClean, reduction = "umap")
Seurat::FeaturePlot(seuClean, features = c("MS4A1", "CD3D", "percent.mito")) #MS4A1 B cell CD3D tcell
Seurat::FeaturePlot(seuClean, features = c("MS4A1"))
Seurat::FeaturePlot(seuClean, features = c("CD3D")) #I SEE DUPLET BC BLEEDING
Seurat::FeaturePlot(seuClean, features = c("percent.mito"))

```

```{r , echo=FALSE}
seuClean <- Seurat::RunUMAP(seuClean, dims = 1:20, n_neighbors = 30L)
Seurat::DimPlot(seuClean, reduction = "umap")
Seurat::FeaturePlot(seuClean, features = c("MS4A1", "CD3D", "percent.mito")) #MS4A1 B cell CD3D tcell
Seurat::FeaturePlot(seuClean, features = c("MS4A1"))
Seurat::FeaturePlot(seuClean, features = c("CD3D")) #I SEE DUPLET BC BLEEDING
Seurat::FeaturePlot(seuClean, features = c("percent.mito"))

```

## Clustering CellAnnotation

```{r, echo=FALSE}
seuClean <- Seurat::FindNeighbors(seuClean, dims = 1:20)

seuClean <- Seurat::FindClusters(seuClean, resolution = seq(0.1, 1, by=0.1))

head(seuClean@meta.data)

library(clustree)
clustree::clustree(seuClean@meta.data[,grep("RNA_snn_res", colnames(seuClean@meta.data))],
                   prefix = "RNA_snn_res.")

```

```{r, echo=FALSE}
# Boucle de 0.1 à 1 par incréments de 0.1
for (res in seq(0.1, 1, by = 0.1)) {
  # Créer la chaîne de caractères correspondant au group.by
  res_string <- paste0("RNA_snn_res.", res)
  
  # Afficher le plot
  print(DimPlot(seuClean, group.by = res_string, label = TRUE) + 
       ggtitle(paste("Résolution:", res)))
} #3 parait stable dans les trees sur le umap
```

## Doublet Finder


```{r, echo=FALSE}
get.data <- function(Sobj, genes) {
  
  # Extract Meta data
  df.meta <- data.frame(Cluster = Sobj@active.ident, Sobj@meta.data)
  
  # Extract tSNE coordinates if available
  if(!is.null(Sobj@reductions$tsne))
  { 
    df.meta <- cbind(df.meta, Sobj@reductions$tsne@cell.embeddings) 
  }
  
  # Extract UMAP coordinates if available
  if(!is.null(Sobj@reductions$umap))
  { 
    df.meta <- cbind(df.meta, Sobj@reductions$umap@cell.embeddings) 
  }
  # Extract gene expression data for expressed genes
  count.data <- as.matrix(GetAssayData(object = Sobj, slot = "scale.data"))
  genes.red <- rownames(count.data)[rownames(count.data) %in% genes]
  dftotal <-  cbind(df.meta, FetchData(Sobj, genes.red))
  
  return(dftotal)
}
```

```{r, echo=FALSE}
CommonMarkers <- c("CD79A", "CD79B", "MS4A1", "CD3E", "CD3D")
DF <- get.data(seuClean, genes=CommonMarkers)
```

```{r, echo=FALSE}

DF$Supervised_Singlets <- "Keep" 
DF$Supervised_Singlets[(DF$CD3E>0 | DF$CD3D>0) & (DF$MS4A1>0 | DF$CD79A>0 | DF$CD79B>0)] <- "Remove"
seuClean$Supervised_Singlets <- DF$Supervised_Singlets

# Number of supervised B/T doublets:
table(seuClean$Supervised_Singlets)

# Visualize the doublets on the 2D graph
Seurat::DimPlot(seuClean, reduction = "umap", group.by = "Supervised_Singlets")
```

```{r, echo= FALSE}
calculate_doubletScore <- function(seuObj) {
  library(DoubletFinder)
  #seu <- RunUMAP(seu, dims = 1:20)
  
  ## pK Identification (no ground-truth) ---------------------------------------------------------------------------------------
  sweep.res.list <- paramSweep(seuObj, PCs = 1:20, sct = FALSE)
  sweep.stats <- summarizeSweep(sweep.res.list, GT = FALSE)
  bcmvn <- find.pK(sweep.stats)
  
  
  nExp_poi <- round(0.04*nrow(seuObj@meta.data))  ## Assuming 0.04% doublet formation rate
  
  ## Run DoubletFinder with varying classification stringencies ----------------------------------------------------------------
  seuObj <- doubletFinder(seuObj, PCs = 1:20, pN = 0.25, pK = 0.09, nExp = nExp_poi, reuse.pANN = FALSE, sct = FALSE)
  return(seuObj)
}

seuClean <- calculate_doubletScore(seuClean)
```

```{r , echo=FALSE}
#Save the results into the Seurat Object
seuClean$Doublets <- seuClean$DF.classifications_0.25_0.09_84 # VERIF DANS LA VARIABLE
seuCleanDoublet <- subset(x = seuClean, subset= Supervised_Singlets=="Keep" & Doublets=="Singlet")
```

```{r, echo=FALSE}
# Visualize the doublets on the 2D graph
table(seuClean$Doublets)
Seurat::DimPlot(seuClean, reduction = "umap", group.by = "Doublets", cols = c("blue", "orange"))
```


```{r, echo=FALSE}
seuCleanDoublet <- Seurat::NormalizeData(seuCleanDoublet,
                             normalization.method = "LogNormalize",
                             scale.factor = 10000)
seuCleanDoublet <- Seurat::FindVariableFeatures(seuCleanDoublet,selection.method = "vst", nfeatures = 2000)
seuCleanDoublet <- Seurat::ScaleData(seuCleanDoublet, features = rownames(seu))

seuCleanDoublet <- Seurat::RunPCA(seuCleanDoublet)
Seurat::ElbowPlot(seuClean, ndims = 40) + 
  geom_hline(yintercept = 5, color = "red", linetype = "dashed") +
  geom_text(aes(x = 13, y = 5, label = round(2.5, 2)), label="5% cut-in threshold", color = "red", vjust = -1, ) +
geom_hline(yintercept = 2.5, color = "green", linetype = "dashed") +
  geom_text(aes(x = 23, y = 3, label = round(2.5, 2)), label="2.5% cut-in threshold", color = "green", vjust = -1, )

dum <- seuCleanDoublet
# Liste des valeurs pour min.dist
min_dist_values <- c(0.1, 0.2, 0.3, 0.4, 0.5)
# Sauvegarder l'identité originale pour pouvoir la restaurer ensuite
original_idents <- Idents(dum)
# Appliquer une identité unique à tous les points pour un affichage uniforme
Idents(dum) <- "All"
# Boucle pour chaque valeur de min.dist
for (dist in min_dist_values) {
  # Appliquer RunUMAP avec la valeur actuelle de min.dist
  dum <- Seurat::RunUMAP(dum, dims = 1:20, min.dist = dist)
  
  # Afficher le DimPlot avec la réduction UMAP
  print(Seurat::DimPlot(dum, reduction = "umap", cols = "orange") + ggtitle(paste("min.dist =", dist)))
}


seuCleanDoublet <- Seurat::RunUMAP(seuCleanDoublet, dims = 1:20)

seuCleanDoublet <- Seurat::FindNeighbors(seuCleanDoublet, dims = 1:20)
seuCleanDoublet <- Seurat::FindClusters(seuCleanDoublet, resolution = seq(0.1, 1, by=0.1))

for (res in seq(0.1, 1, by = 0.1)) {
  # Créer la chaîne de caractères correspondant au group.by
  res_string <- paste0("RNA_snn_res.", res)
  
  # Afficher le plot
  print(DimPlot(seuCleanDoublet, group.by = res_string, label = TRUE) + 
       ggtitle(paste("Résolution:", res)))
}

clustree::clustree(seuCleanDoublet@meta.data[,grep("RNA_snn_res", colnames(seuCleanDoublet@meta.data))],
                   prefix = "RNA_snn_res.")
```



## Cell Annotation

```{r, echo=FALSE}
library(Seurat)
library(celldex)
library(SingleR)
```

```{r, echo=FALSE}
seuCleanDoublet <- Seurat::SetIdent(seuCleanDoublet, value = seuCleanDoublet$RNA_snn_res.0.4)
```

```{r, echo=FALSE}
DefaultAssay(seuCleanDoublet) <- "RNA"
Seurat::FeaturePlot(seuCleanDoublet, c("MS4A1", "CD3D"))

```
```{r, echo=FALSE}
markers_cluster0 <- FindMarkers(seuCleanDoublet, ident.1 = 9)
# Filtrer pour ne garder que les gènes avec un p-valeur ajusté < 0.05
significant_genes <- markers_cluster0[markers_cluster0$p_val_adj < 0.001, ]
gene_names <- rownames(significant_genes)
print(gene_names)

```

```{r, echo=FALSE}
string <- "WWTR1"
if (string %in% gene_names) {
  print("true")
} else {
  print("false")
}
```

```{r, echo=FALSE}
tumorSuppresor <- c("DLC1","LOX","PTPRG")

Seurat::FeaturePlot(seuCleanDoublet, tumorSuppresor, ncol=2)
Seurat::VlnPlot(seuCleanDoublet,
                features = tumorSuppresor,
                ncol = 2)
```

```{r, echo=FALSE}

protoOncogene <- c("JUNB", "FOS", "COL5A2", "TGFBR3")

Seurat::FeaturePlot(seuCleanDoublet, protoOncogene, ncol=2)
Seurat::VlnPlot(seuCleanDoublet,
                features = protoOncogene,
                ncol = 2)
```


```{r, echo=FALSE}
tcell_genes <- c("IL7R", "TRAC", "CD3D", "CD3E")
bcell_genes <- c("MS4A1", "CD79A", "CD79B","CD19")

Seurat::FeaturePlot(seuCleanDoublet, tcell_genes, ncol=2)
Seurat::VlnPlot(seuCleanDoublet,
                features = tcell_genes,
                ncol = 2)

```

```{r, echo=FALSE}
Seurat::FeaturePlot(seuCleanDoublet, bcell_genes, ncol=2)
Seurat::VlnPlot(seuCleanDoublet,
                features = bcell_genes,
                ncol = 2)
```


```{r, echo=FALSE}
seuCleanDoublet <- Seurat::AddModuleScore(seuCleanDoublet,
                              features = list(bcell_genes),
                              name = "bcell_genes")
seuCleanDoublet <- Seurat::AddModuleScore(seuCleanDoublet,
                              features = list(tcell_genes),
                              name = "tcell_genes")

seuCleanDoublet <- Seurat::AddModuleScore(seuCleanDoublet,
                              features = list(protoOncogene),
                              name = "protoOncogene")
seuCleanDoublet <- Seurat::AddModuleScore(seuCleanDoublet,
                              features = list(tumorSuppresor),
                              name = "tumorSuppresor")
```

```{r, echo=FALSE}
Seurat::FeaturePlot(seuCleanDoublet, "bcell_genes1")
Seurat::FeaturePlot(seuCleanDoublet, "tcell_genes1")
Seurat::FeaturePlot(seuCleanDoublet, "protoOncogene1")
Seurat::FeaturePlot(seuCleanDoublet, "tumorSuppresor1")
```

### Cycling phase

```{r, echo=FALSE}
s.genes <- Seurat::cc.genes.updated.2019$s.genes
g2m.genes <- Seurat::cc.genes.updated.2019$g2m.genes
```

```{r, echo=FALSE}
seuCleanDoublet <- Seurat::CellCycleScoring(seuCleanDoublet, s.features = s.genes, g2m.features = g2m.genes)
```

```{r, echo=FALSE}
Seurat::DimPlot(seuCleanDoublet, group.by = "Phase")
```

### SingleR

#### BlueprintEncodeData

```{r, echo=FALSE}
ref <- celldex::BlueprintEncodeData()

# Let's take a look at the cell types present in the refrence
table(ref$label.main)

# Let's select from the reference the populations of cells likely to be present in our dataset 
ref_reduce <- ref #[,ref$label.main %in% c("B-cells","CD4+ T-cells","CD8+ T-cells","Monocytes","NK cells","Epithelial cells","HSC","Eosinophils","DC")] #CHANGE
```

```{r, echo=FALSE}
library(pheatmap)
seu_SingleR <- SingleR(test=GetAssayData(seuCleanDoublet, assay = "RNA", slot = "data"),
                           ref=ref_reduce, labels=ref_reduce$label.fine)
head(seu_SingleR)
SingleR::plotScoreHeatmap(seu_SingleR)
```

```{r, echo=FALSE}
seuCleanDoublet$SingleR_annot <- seu_SingleR$labels
dittoSeq::dittoDimPlot(seuCleanDoublet, "SingleR_annot", size = 0.7)
dittoSeq::dittoBarPlot(seuCleanDoublet, var = "SingleR_annot", group.by = "RNA_snn_res.0.4")
```

### CHETAH

```{r, echo=FALSE}
library(CHETAH)

load("CHETAH_TME_reference.Rdata") 

parotid10x.sce <- CHETAH::CHETAHclassifier(input=as.SingleCellExperiment(seuCleanDoublet), ref_cells = reference, thresh = 0.05)

PlotCHETAH(parotid10x.sce, redD="UMAP")

PlotCHETAH(parotid10x.sce, redD="UMAP", tree = FALSE)

# Let's visualize the cell annotation
print(table(parotid10x.sce$celltype_CHETAH))

seuCleanDoublet$CHETAH_annot <- as.character(parotid10x.sce$celltype_CHETAH)
```

## Cell_CrossTalks

### SingleR

```{r, echo=FALSE}
library(CellChat)  #BiocManager::install("ComplexHeatmap")
#devtools::install_github("sqjin/CellChat")
library(SingleR)
```

```{r, echo=FALSE}
# Select only T, B and plasmacells and fibroplaste car cancer

sce.Parotid10x <- subset(seuCleanDoublet, subset= SingleR_annot=="Monocytes" | SingleR_annot=="CD4+ Tcm" | SingleR_annot=="Macrophages" | SingleR_annot=="CD8+ Tcm" | SingleR_annot=="Epithelial cells" | SingleR_annot=="Tregs" | SingleR_annot=="Plasma cells" | SingleR_annot=="Fibroblasts" | SingleR_annot=="Class-switched memory B-cells" | SingleR_annot=="CD4+ Tem" | SingleR_annot=="NK cells" | SingleR_annot=="DC" | SingleR_annot=="CD8+ Tem" | SingleR_annot=="Adipocytes" | SingleR_annot=="CLP" )
  #subset(seuCleanDoublet, subset= SingleR_annot!="CD4+ T-cells" SingleR_annot!="CD8+ T-cells" |SingleR_annot!="Endothelial cells"| SingleR_annot!="MEP" | SingleR_annot!="MPP" | SingleR_annot!="mv Endothelial cells" | SingleR_annot!="Memory B-cells" | SingleR_annot!="naive B-cells" | SingleR_annot!="CMP" | SingleR_annot!="Myocytes" | SingleR_annot!="Macrophages M1" | SingleR_annot!="HSC" | SingleR_annot!="Keratinocytes" | SingleR_annot!="Mesangial cells") #seuCleanDoublet #subset(seuCleanDoublet, subset= SingleR_annot=="CD4+ T-cells" | SingleR_annot=="CD4+ Tcm" | SingleR_annot=="CD8+ T-cells" | SingleR_annot=="CD8+ Tcm" | SingleR_annot=="Memory B-cells" | SingleR_annot=="naive B-cells" | SingleR_annot=="Plasma cells" | SingleR_annot=="Fibroblasts")




```

```{r, echo=FALSE}
CellChatDB <- CellChatDB.human  # Note: use CellChatDB.mouse if running on mouse data
#showDatabaseCategory(CellChatDB)
```

```{r, echo=FALSE}
# Create the CellChat object and specify the name of the column in the meta.data slot containing the cell group or annotation
cellchatParotid10x <- createCellChat(object = sce.Parotid10x, group.by = "SingleR_annot")

# set the used database in the object
cellchatParotid10x@DB <- CellChatDB

cellchatParotid10x <- subsetData(cellchatParotid10x) # subset the expression data of signaling genes for saving computation cost

# Identify over-expressed ligands or receptors
cellchatParotid10x <- identifyOverExpressedGenes(cellchatParotid10x, group.by = "SingleR_annot")

# Identify over-expressed ligand-receptor interactions if either ligand or receptor is over-expressed.

cellchatParotid10x <- identifyOverExpressedInteractions(cellchatParotid10x)
cellchatParotid10x <- projectData(cellchatParotid10x, PPI.human)
cellchatParotid10x <- computeCommunProb(cellchatParotid10x, raw.use = TRUE)

# Filter out the cell-cell communication if there are only few number of cells in certain cell groups
cellchatParotid10x <- filterCommunication(cellchatParotid10x, min.cells = 10)

# Compute the communication probability and infer cellular communication network
cellchatParotid10x <- computeCommunProbPathway(cellchatParotid10x)

# Calculate the aggregated cell-cell communication network
cellchatParotid10x <- aggregateNet(cellchatParotid10x)

groupSize <- as.numeric(table(cellchatParotid10x@idents))
par(mfrow = c(1,1), xpd=TRUE)

# Let's visualize the interactions (crosstalks) within our cells
netVisual_circle(cellchatParotid10x@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Parotid 10x\nNumber of interactions")
```

Due to the complicated cell-cell communication network, we can examine the signaling sent from CD4 T-cells and the other cell groups

```{r, echo=FALSE}
mat <- cellchatParotid10x@net$weight
par(mfrow = c(1, 1), xpd = TRUE)

# Taille de `mat` pour obtenir le nombre de lignes
n <- nrow(mat)

for (i in 1:n) {
  # Créer une matrice temporaire pour chaque ligne i
  temp_mat <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  
  # Copier les valeurs de la ligne i de `mat` dans `temp_mat`
  temp_mat[i, ] <- mat[i, ]
  
  # Définir un titre dynamique pour le graphique en fonction de l'indice de ligne
  titre <- paste("Graphique pour ligne", rownames(mat)[i], i)
  
  # Visualiser le graphique
  netVisual_circle(temp_mat, vertex.weight = groupSize, 
                   weight.scale = TRUE, edge.weight.max = max(mat), title.name = titre)
}

```

Then take a look at the significant Ligand/Receptor pairs:

```{r, echo=FALSE}
n <- nrow(mat)

for (i in 1:n) {
  j <- i + 1
  titre <- paste("10x MALT \n  Graphique pour ligne", rownames(mat)[i], i)
  tryCatch({
    interaction <- netVisual_bubble(
      cellchatParotid10x, 
      sources.use = c(i:j),
      #targets.use = c(1:n),  # Inclure tous les groupes comme cibles, si nécessaire
      remove.isolate = TRUE,
      title.name = titre,
      #return.data = T
    )
    print(interaction)
  }, error = function(e) {
    # Affiche un message d'avertissement et passe à l'itération suivante
    print(paste("Aucune interaction détectée pour le groupe", rownames(mat)[i], i))
  })
}
```

We can plot the gene expression distribution of signaling genes related to L-R pairs or signaling pathway using a Seurat wrapper function `plotGeneExpression`.

```{r, echo=FALSE}
plotGeneExpression(cellchatParotid10x, signaling = "COLLAGEN") #regarde dans pathways pour mettre le bon signaling DEMANDE A CHATON ON VEUT ETUDIE QUOI

plotGeneExpression(cellchatParotid10x, signaling = "TGFb")
```

```{r, echo=FALSE}
for (i in 1:n) {
    titre <- paste("10x MALT \n Graphique pour ligne", rownames(mat)[10], 10," <- ", rownames(mat)[i])
    tryCatch({
      # Utiliser uniquement la cellule 10 en tant que source
      interaction <- netVisual_bubble(
        cellchatParotid10x, 
        sources.use = 10,  # Cellule 14 comme source unique
        targets.use = i,   # Cible est chaque cellule "i" dans la boucle
        remove.isolate = TRUE,
        title.name = titre,
        font.size = 5
      )
      print(interaction)
    }, error = function(e) {
      # Affiche un message d'avertissement si aucune interaction n'est détectée
      print(paste("Aucune interaction détectée pour le groupe", rownames(mat)[i]))
    })
}
```

## Cell Trajectory

```{r, echo=FALSE}
# This seurat object has two vectors in the meta.data cointining important information 
table(seuCleanDoublet$seurat_clusters)

table(seuCleanDoublet$Supervised_Singlets)
```

```{r, echo=FALSE}
library("RColorBrewer")
library("scater") #BiocManager::install("scater")
library("TSCAN")  #BiocManager::install("TSCAN")
library("destiny") #BiocManager::install("destiny")
library("slingshot") #BiocManager::install("slingshot")
library("Seurat")
```

For this example here we are excluding from the analysis T-cells, Monocytes and cells present at low percentage.

```{r, echo=FALSE}
# Subset the seurat object and select only B-cells

seuCleanDoublet_Traj <- subset(x = seuCleanDoublet, subset= SingleR_annot=="Fibroblasts" | SingleR_annot=="Monocytes" | SingleR_annot=="Macrophages" | SingleR_annot=="DC" | SingleR_annot=="HSC") 
```

We could now extract the counts and meta.data to create the SingleCellExperiment object

```{r, echo=FALSE}
counts <- seuCleanDoublet_Traj@assays$RNA@counts 
metadata <- seuCleanDoublet_Traj@meta.data

# Set up metadata as desired for aggregation and DE analysis
Idents(seuCleanDoublet_Traj) <- seuCleanDoublet_Traj$SingleR_annot
metadata$cluster_id <- factor(seuCleanDoublet_Traj@active.ident)

# Create single cell experiment object
sce <- SingleCellExperiment(assays = list(counts = counts), 
                            colData = metadata)
reducedDims(sce) <- list(PCA=seuCleanDoublet_Traj[["pca"]]@cell.embeddings[, 1:20],
                         UMAP=seuCleanDoublet_Traj[["umap"]]@cell.embeddings)
```

So, everything is ready for the trajectory inference using slingshot R package

```{r, echo=FALSE}
  gastric_SCE <- slingshot(sce, clusterLabels = 'SingleR_annot', reducedDim = 'PCA')
  

  rd <- reducedDims(gastric_SCE)$UMAP

  lin1 <- getLineages(rd, seuCleanDoublet_Traj$SingleR_annot)
  
  # Command for visualization purpose
  
  seuCleanDoublet_Traj$SingleR_annot <- as.factor(seuCleanDoublet_Traj$SingleR_annot)
  plot(rd, col = brewer.pal(3,"Set1")[seuCleanDoublet_Traj$SingleR_annot], asp = 1, pch = 16)
  lines(SlingshotDataSet(lin1), lwd = 3, col = 'black')
  Seurat::DimPlot(seuCleanDoublet_Traj, reduction = "umap") #, group.by = "Supervised_Singlets"
```
